Future challenges for Research Software Engineering

### 2025-05-17 future_challenges_for_RSEng.tex
swap words around
split sentence

The specialist skills that RSEs provide take a significant amount of time to develop and maintain.
We would, therefore, not expect domain researchers to also become experts in a range of
advanced RSE topics. 
At the same time, we do recognise that the application of core software development best practices
to enhance and address problems with research codebases, that can often make up the bulk of
current RSE work, will change. As noted above, these are skills that we
would expect any researcher who writes code to have in 10 years time. 
The rapid development of AI means that researchers are likely to have access to tooling that can assist
them in writing robust, sustainable code, lowering the barrier to developing code, or at least changing the
required skill set to some extent.
They can also be expected to have an understanding of frameworks to support testing, packaging and
deployment of their code as well as integration with central domain specific infrastructures.
Again, all these skills will, we anticipate, be gained as part of training infrastructure that will
provide domain scientists with software development expertise as part of their core scientific training.

@@

It is already difficult to structure the spectrum of RSE tasks and responsibilities today given the spectrum of research software \cite{hasselbring2024}
and it is even more challenging to structure the aspects that will become relevant for the RSE community in the future.
Nevertheless, we decided to consider:
\begin{itemize}
\item The evolution of the term RSE
\item Development of the RSE community
\item What the impact of AI will be to the work of RSEs
\item The impact of complexity and usability in research software
\item Ethics and social considerations
\item How education is reshaped by the digital transformation
\item Changes to academic institutions
\item Impact of alternate career paths for RSEs and other RTPs
\end{itemize}

>>>>

The specialist skills that RSEs provide take a significant amount of time to develop and maintain.
Therefore, we do not expect domain researchers to also become experts in a range of
advanced RSE topics. 
At the same time, we do recognise that the application of core software development best practices
to enhance and address problems with research codebases, that can often make up the bulk of
current RSE work, will change. As noted above, these are skills that we
would expect any researcher who writes code to have in 10 years time. 
The rapid development of AI means that researchers are likely to have access to tooling that can assist
them in writing robust, sustainable code, lowering the barrier to developing code, or at least changing the
required skill set to some extent.
They can also be expected to have an understanding of frameworks to support testing, packaging and
deployment of their code as well as integration with central domain specific infrastructures.
Again, all these skills will, we anticipate, be gained as part of training infrastructure that will
provide domain scientists with software development expertise as part of their core scientific training.

@@

It is already difficult to structure the spectrum of RSE tasks and responsibilities today given the spectrum of research software \cite{hasselbring2024}.
It is even more challenging to structure the aspects that will become relevant for the RSE community in the future.
Nevertheless, we decided to consider:
\begin{itemize}
\item The evolution of the term RSE
\item Development of the RSE community
\item What the impact of AI will be to the work of RSEs
\item The impact of complexity and usability in research software
\item Ethics and social considerations
\item How education is reshaped by the digital transformation
\item Changes to academic institutions
\item Impact of alternate career paths for RSEs and other RTPs
\end{itemize}
